---
title: "#DuboisChallenge Tweets"
author: "Antony Rono"
date: 2021-06-19
output: html_output
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(lubridate)
library(plotly)
library(magrittr)
library(ggthemes)

options(scipen = 99)
theme_set(theme_classic())

```

# Load the weekly Data

```{r Load}

tt <- tt_load("2021-06-15")

tweets <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-15/tweets.csv')

```

# Cleaning the data

## Dates and Usernames

```{r data cleaning}

tweets_cleaned <- tweets %>% 
  
  separate(datetime, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
  
  mutate(date = ymd(date),
         
         datetime = ymd_hms(datetime),
         
         hour = str_extract(floor_date(datetime, "hour"), "[0-9]{2}\\:[0-9]{2}"),
         
         week_day = wday(date, label = TRUE, abbr = FALSE),
         
         username = paste0("@", username)
         
         ) %>% 
  
  select(1:3, hour, week_day, everything()) 

```

## Primary Data to be used in maps

```{r map data}

tweets_map <- tweets_cleaned %>% 
  
  filter(!is.na(long), !is.na(lat)) %>% 
  
  rowwise() %>% 
  
  mutate(activity = sum(c_across(ends_with("count")))) %>% 
  
  ungroup() %>% 
  
  select(location,long, lat, activity) %>% 
  
  rename(lon = long) 


```

## Top 10 tweets with highest interactions

```{r tweets with highest interactions}


tweets_highest_interactions <- tweets_cleaned %>% 
  
  rowwise() %>% 
  
  mutate(activity = sum(c_across(ends_with("count"))),
         
         url = paste0("<a href=\"", url, "\">Link</a>"),
         
         date = format(date, "%d %b")
         ) %>% 
  
  arrange(desc(activity)) %>% 
  
  head(10)

tweets_highest_interactions %<>% 
  
  #mutate(activity = paste0("Username:", username, "\nRetweets:", retweet_count, "\nLikes:", like_count , "\nQuotes", quote_count))

  mutate(activity = paste0(username, " on ", date, ":\n", retweet_count, " Retweets, ", like_count, " Likes, ", quote_count , " Quotes"))

```

# Visualization

We explore 3 approaches of plotting maps:

1. Using Maps package

2. Using SF & ggplot2 Packages (borders function)

3. Using SF & ggplot2 Packages (geom_path function)

4. Highcharter map (See the "2021_06_15_tidy_tuesday_dashboard,Rmd" for this)


Note: We've used the 4th approach to generate the plot


#### Using Maps Package

```{r map of activity - maps package}

library(maps)

map("world", fill=TRUE, col="white", bg="lightblue", ylim=c(-60, 90), mar=c(0,0,0,0)) 

points(tweets_map$lon, tweets_map$lat, col = "red", pch = 16)
  
  
```

#### Using sf Package - Approach 1: borders function

```{r map of activity - sf package: approach 1}
library(ggrepel)
library(sf)
library(ggtext)


tweets_map_sf <- st_as_sf(tweets_map, coords = c("lon", "lat"), crs = 4326)  

world_map <- map_data("world")


ggplot() +
  
  borders("world", colour="gray50", fill="white")+ # create a layer of borders
  
  geom_sf(data = tweets_map_sf, aes(color = "red", size =activity) ) +
  
  geom_text_repel(aes(x = long , y = lat, label = activity),data = tweets_highest_interactions,
                   size= 3,
                   fontface = 'bold',
                   color = "violetred",
                   segment.colour = "violetred",
                  ) +
  
  theme_map() +
  
  theme(plot.background = element_rect(fill = '#dbc4b0'),
        legend.position = "none",
        plot.title = element_text(size=18, color = "forestgreen", face = "bold"),
        plot.subtitle = element_markdown(lineheight = 1.1, size = 11)
        ) +
  
  labs(title = "#DUBOISCHALLENGE TWEETS - 2021 .",
       
        subtitle = "A total of <span style='color:red;'>**445**</span> tweets were tweeted with the <span style='color:forestgreen;'>**#DuBoisChallenge**</span> tag.<br>
        This chart shows the distribution of the tweets across the globe, highlighting the <span style='color:violetred;'>**top 10 tweets with the highest activity**</span>",
        fill = NULL, color = NULL, y = NULL, x = "Average record time for a 3-lap course (seconds)",
       
        caption = "source :Anthony Starks, Allen Hillery & Sekou Tyler. | viz: Antony Rono"
       
       )
  

```

#### Using sf Package - geom_path

```{r map of activity - sf package: appraoch 2}

# library(ggrepel)
# library(sf)
# library(ggtext)
# 
# 
# tweets_map_sf <- st_as_sf(tweets_map, coords = c("lon", "lat"), crs = 4326)  
# 
# world_map <- map_data("world")


ggplot() +
  
  geom_path(aes(x=long, y=lat, group=group), data = world_map) +
  
  scale_y_continuous(breaks=(-2:2) * 30) +
  
  scale_x_continuous(breaks=(-4:4) * 45) +
  
  geom_sf(data = tweets_map_sf, aes(color = "darkred", size = activity)) +
  
  geom_text_repel(aes(x = long , y = lat, label = activity),data = tweets_highest_interactions,
                   size= 3.5,
                   fontface = 'bold',
                   color = "violetred",
                   segment.colour = "violetred",
                  ) +
  
  theme_map() +
  
  theme(plot.background = element_rect(fill = '#dbc4b0'),
        legend.position = "none",
        plot.title = element_text(size=18, color = "forestgreen", face = "bold"),
        plot.subtitle = element_markdown(lineheight = 1.1, size = 12)
        ) +
  
  labs(title = "#DUBOISCHALLENGE TWEETS - 2021.",
       
        subtitle = "A total of <span style='color:red;'>**445**</span> tweets were tweeted with the <span style='color:forestgreen;'>**#DuBoisChallenge**</span> tag.<br>
        This chart shows the distribution of the tweets across the globe, highlighting the <span style='color:violetred;'>**top 10 tweets with the highest interactions**</span>",
        fill = NULL, color = NULL, y = NULL, x = "Average record time for a 3-lap course (seconds)",
       
        caption = "source :Anthony Starks, Allen Hillery & Sekou Tyler. | viz: Antony Rono"
       
       )

```

```{r save plot}
ggsave("plots/DuBoisChallenge_tweets.png", 
       bg = "transparent", 
       width = 14, 
       height = 8, 
       dpi = 600)
```


