---
title: "Animal Rescues: 2009-2020"
author: "Antony Rono"
date: 2021-07-04
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(flexdashboard)
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(janitor)
library(lubridate)
library(magrittr)
library(plotly)

theme_set(theme_light())
```

```{r Load, include=FALSE}

raw_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-29/animal_rescues.csv')

```

```{r Cleaning data, include=FALSE}

animal_rescues <- raw_data %>% 
  
  mutate(date_time_of_call = dmy_hm(date_time_of_call),
         
         animal_group_parent = str_to_sentence(animal_group_parent),
         
         year_month = as.Date(floor_date(date_time_of_call, "month")),
         
         month = month(date_time_of_call, label = TRUE, abbr = FALSE),
         
         year = year(date_time_of_call),
         
         incident_notional_cost = parse_number(incident_notional_cost)
         
         ) %>% 
  
  filter(year < 2021)



```

```{r function to plot bar charts, include=FALSE}

# Make a palette by drawing colors from `RColorBrewer`

custom_palette <- c(RColorBrewer::brewer.pal(8, "Dark2"),
                    RColorBrewer::brewer.pal(9, "Set1")
                    
                    )


plot_categorical_bar <- function(tbl, col,wt = NULL, reorder = TRUE, nlevel = 8, coord_flip = TRUE){
  
  lumped_data <- tbl %>% 
    
    filter(!is.na({{col}})) %>% 
  
    mutate({{col}} :=fct_lump({{col}}, nlevel)) %>% 
  
    count({{col}},wt = {{wt}}, name = "Count")
  
  if(reorder){
    
    lumped_data %<>%  
      
      mutate({{col}} := fct_reorder({{col}}, Count))
    
  }
  
  p <- lumped_data %>% 
  
    ggplot(aes({{col}}, Count, fill = {{col}})) +
  
    geom_col() +
  
    theme(legend.position = "none") +
  
    scale_y_continuous(expand = c(0,0),labels = comma)+
    
    scale_fill_manual(values = custom_palette) +
  
    theme(legend.position = "none") 
  
  
  if(coord_flip){
    
    p + coord_flip() +
      
      theme(axis.title.y = element_blank())
    
  }else{
    
    p +
      
      theme( axis.title.x = element_blank())
  
    }
  
}


```




Overall {data-icon="fa-map-marker-alt"}
=======================================================================
  
Row {data-height=200}
-----------------------------------------------------------------------
  
### Total Rescues
  
```{r total rescues}

valueBox(value = comma(nrow(animal_rescues)), caption = "Total Number of Rescues", icon = "fa-hashtag", color = "green")

```

### Total Notional Cost

```{r total cost}

cost <- paste0("£",comma(sum(animal_rescues$incident_notional_cost, na.rm = TRUE)))

valueBox(value = cost, caption = "Total Notional Cost", icon = "fa-coins", color = "blue")

```

### Most Rescued Animal

```{r most rescued animal}

animal <- animal_rescues %>% 
  
  count(animal_group_parent, sort = TRUE) %>% 
  
  head(1)

value <- paste0(animal$animal_group_parent, " (",comma(animal$n), ")", " Rescues" )

valueBox(value = value, caption = "Most Rescued Animal", icon = "fa-users", color = "coral")

```

Row {data-height=800}
-----------------------------------------------------------------------
  
### Rescues by Animal Group Parent
  
```{r}

p <- animal_rescues %>% 
  
  plot_categorical_bar(animal_group_parent)

ggplotly(p, tooltip = c("x", "y"))

```

### Rescues by Property Category

```{r}

p <- animal_rescues %>% 
  
  plot_categorical_bar(property_category, nlevel = 5)

ggplotly(p, tooltip = c("x", "y"))

```

### Rescues by Special Services Category

```{r}

p <- animal_rescues %>% 
  
  plot_categorical_bar(special_service_type_category)

ggplotly(p, tooltip = c("x", "y"))

```


Rescues by Month {data-icon="fa-twitter"}
=======================================================================

Row {data-height=200}
-----------------------------------------------------------------------

### Month with the Highest Number of Animal Rescue

```{r user with most activity}

month_dt <- animal_rescues %>%
  
  count(month, sort = TRUE) %>% head(1)
  
value <- paste0(month_dt$month, "(",comma(month_dt$n), " Rescues", ")" )

valueBox(value = value, caption = "Month with Most Rescues", icon = "fa-calendar", color = "tomato")


```

### Month with the Highest Cost of Animal Rescue

```{r date with most activity}

month_dt <- animal_rescues %>%
  
  count(month, wt = incident_notional_cost, sort = TRUE) %>% head(1)
  
value <- paste0( month_dt$month, "(","£", comma(month_dt$n), ")" )

valueBox(value = value, caption = "Month with Highest Rescue Cost", icon = "fa-pound-sign", color = "blueviolet")


```

Row {data-height=800}
-----------------------------------------------------------------------

### Number of Rescues by Month

```{r}

p <- animal_rescues %>% 
  
  plot_categorical_bar(month,  reorder = FALSE, nlevel = 12, coord_flip = FALSE) +
  
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

ggplotly(p, tooltip = c("x", "y"))

```

### Incident Notinal Cost by Month

```{r}

p <- animal_rescues %>% 
  
  plot_categorical_bar(month, wt = incident_notional_cost, reorder = FALSE, nlevel = 12, coord_flip = FALSE) +
  
  scale_y_continuous(labels = scales::unit_format(unit = "K", scale = 1e-3,prefix = "£"),
                     
                     expand = c(0,0)) +
  
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  
  labs(y = "Incident Notional Cost")

ggplotly(p, tooltip = c("x", "y"))
```


Trends {data-icon="fa-chart-line" data-orientation=columns}
=======================================================================

Column{.tabset}
-----------------------------------------------------------------------

### Monthly Trend in Animal Rescues

```{r}

p <- animal_rescues %>% 
  
  count(year_month) %>% 
  
  ggplot(aes(year_month, n))+
    
  geom_col(fill = "steelblue")+
  
  geom_line(color = "red")+
  
  scale_y_continuous(expand = c(0,0))+
  
  scale_x_date(expand = c(0,0),labels = date_format("%b %y"), breaks = date_breaks("years"))+
  
  theme(legend.position = "none") +
  
    labs(x = "",
       
       y = "Count")

ggplotly(p, tooltip = c("x", "y"))
```

### Yearly Trend in Animal Rescues

```{r}

p <- animal_rescues %>% 
  
  count(year) %>% 
  
  filter(year < 2021) %>% 
  
  ggplot(aes(year, n))+
    
  geom_col(aes(fill = as.character(year)))+
  
  geom_line(aes(group = 0), color = "coral", size = 1.2)+
  
  scale_y_continuous(expand = c(0,0))+
  
  scale_x_continuous(expand = c(0,0))+
  
  scale_fill_manual(values = custom_palette) +
  
  theme(legend.position = "none") +
  
    labs(x = "",
       
       y = "Count")

ggplotly(p, tooltip = c("x", "y"))
```

Column{.tabset}
-----------------------------------------------------------------------

### Monthly Trend in Incident Notional Cost

```{r}

p <- animal_rescues %>% 
  
  filter(!is.na(incident_notional_cost)) %>% 
  
  count(year_month, wt = incident_notional_cost) %>% 
  
  ggplot(aes(year_month, n))+
    
  geom_col(fill = "coral")+
  
  geom_line(color = "steelblue")+
  
  scale_y_continuous(expand = c(0,0))+
  
  scale_x_date(expand = c(0,0),labels = date_format("%b %y"), breaks = date_breaks("years"))+
  
  theme(legend.position = "none") +
  
  labs(x = "",
       
       y = "Incident Notional Cost")


ggplotly(p, tooltip = c("x", "y"))
```

### Yearly Trend in Incident Notional Cost

```{r}

p <- animal_rescues %>% 
  
  filter(!is.na(incident_notional_cost)) %>% 
  
  count(year, wt = incident_notional_cost) %>% 
  
  ggplot(aes(year, n))+
    
  geom_col(aes(fill = as.character(year)))+
  
  geom_line(aes(group = 1), color = "coral", size = 1.2)+
  
  scale_y_continuous(expand = c(0,0), labels = comma)+
  
  scale_x_continuous(expand = c(0,0))+
  
  scale_fill_manual(values = custom_palette) +
  
  theme(legend.position = "none") +
  
  labs(x = "",
       
       y = "Incident Notional Cost")

ggplotly(p, tooltip = c("x", "y"))

```
