---
title: "Mario Kart World Records"
author: "Antony Rono"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(tidytuesdayR)
library(janitor)
library(scales)
library(lubridate)
library(plotly)
library(hrbrthemes)

options(scipen = 99)
theme_set(theme_light())

```


```{r Load data, include=FALSE}

#tt <- tt_load("2021-05-25")

records <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/records.csv')
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/drivers.csv')

```


Trends in World Records {data-icon="fa-chart-line"}
=======================================================================

Column
-----------------------------------------------------------------------

### Overall

```{r trend in world record - overall}

p <- records %>% 
  
  mutate(year = year(date)) %>% 
  
  group_by(year, type, track) %>% 
  
  summarise(record_time = min(time)) %>% 
  
  ungroup() %>% 
  
  ggplot(aes(year, record_time, color = type)) +
  
  geom_line() +
  
  geom_point() +
  
  facet_wrap(~track) +
  
  theme_bw()+
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1),
        
        strip.background = element_blank(),
        
        panel.border = element_rect(colour = "black", fill = NA),
        
        strip.text = element_text(size = 8)) 
p
ggplotly(p)

```

Column {.tabset}
-----------------------------------------------------------------------

### Without Shortcuts

```{r trend in world record - without shortcut}


p <- records %>% 
  
  filter(shortcut == "No") %>% 
  
  mutate(year = year(date)) %>% 
  
  group_by(year, type, track) %>% 
  
  summarise(record_time = min(time)) %>% 
  
  ungroup() %>% 
  
  ggplot(aes(year, record_time, color = type)) +
  
  geom_line() +
  
  geom_point() +
  
  facet_wrap(~track) +
  
  theme_bw()+
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1),
        
        strip.background = element_blank(),
        
        panel.border = element_rect(colour = "black", fill = NA),
        
        strip.text = element_text(size = 8)) 

ggplotly(p)

```

### With Shortcut

```{r trend in world record - with shortcut}

p <- records %>% 
  
  filter(shortcut == "Yes") %>% 
  
  mutate(year = year(date)) %>% 
  
  group_by(year, type, track) %>% 
  
  summarise(record_time = min(time)) %>% 
  
  ungroup() %>% 
  
  ggplot(aes(year, record_time, color = type)) +
  
  geom_line() +
  
  geom_point() +
  
  facet_wrap(~track) +
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1)) +
  
  theme_bw()+
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1),
        
        strip.background = element_blank(),
        
        panel.border = element_rect(colour = "black", fill = NA),
        
        strip.text = element_text(size = 8)) 



ggplotly(p)
```

Fastest Tracks {data-icon="fa-space-shuttle"}
=======================================================================

Column (data-width=750)
-----------------------------------------------------------------------

### Fastest Tracks

```{r fastest tracks}

p <- records %>% 
  
  mutate(shortcut = case_when(
    
    shortcut == "No" ~ "Without Shortcut",
    
    shortcut == "Yes" ~ "With Shortcut",
    
    TRUE ~ shortcut
    
    ),
  
    shortcut = fct_relevel(shortcut, "Without Shortcut")
  
  ) %>% 
  
  group_by(track, type, shortcut) %>% 
  
  summarise(min_time = min(time)) %>% 
  
  ungroup() %>% 
  
  group_by(track, shortcut) %>% 
  
  mutate(min = min(min_time)) %>% 
  
  arrange(min, .by_group = TRUE) %>% 
  
  ungroup() %>% 
  
  mutate(track = fct_reorder(track, min)) %>% 
  
  ggplot(aes(min_time, track, fill = type))+
  
  geom_col(position = "dodge") +
  
  scale_x_continuous(expand = c(0,0))+
  
  facet_wrap(~shortcut) +
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1),
        
        #strip.background = element_blank(),
        
        panel.border = element_rect(colour = "black", fill = NA),
        
        strip.text = element_text(size = 8)) 
 

ggplotly(p) 

```

Column (data-width=250)
-----------------------------------------------------------------------

### Observations

- Mario Raceway is the fastest track without shortcut (Time 27.62 seconds)

- Wario Stadium Fastest without Shortcut (Time 14.59 seconds)


Improvement in World Records {data-icon="fa-sort-up"}
=======================================================================

Column
-----------------------------------------------------------------------

### Overall

```{r improvements in word record - overall}


p <- records %>% 
  
  group_by(track, type) %>% 
  
  summarise(change = max(time) - min(time),
         
         pct_change = change/max(time)) %>% 
  
  mutate(total = 1 -pct_change) %>% 
  
  ungroup() %>% 
  
  select(-change) %>% 
  
  mutate(track = fct_reorder(track, pct_change, .desc = FALSE)) %>% 
  
  ggplot(aes(pct_change, track, fill = type)) +
  
   geom_col()+
  
  #geom_col(aes(x = total, track)
  
  scale_x_continuous(labels = percent,
                     expand = c(0,0)) +
  
    facet_wrap(~type) +
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1),
        
        #strip.background = element_blank(),
        
        panel.border = element_rect(colour = "black", fill = NA),
        
        strip.text = element_text(size = 10)) 

ggplotly(p)

```

Column {.tabset}
-----------------------------------------------------------------------

### Long Track

```{r improvements in word record - long track}

p <- records %>% 
  
  filter(shortcut == "No") %>% 
  
  group_by(track, type) %>% 
  
  summarise(change = max(time) - min(time),
         
         pct_change = change/max(time)) %>% 
  
  mutate(total = 1 -pct_change) %>% 
  
  ungroup() %>% 
  
  select(-change) %>% 
  
  mutate(track = fct_reorder(track, pct_change, .desc = FALSE)) %>% 
  
  ggplot(aes(pct_change, track, fill = type)) +
  
   geom_col()+
  
  scale_x_continuous(labels = percent,
                     expand = c(0,0)) +
  
    facet_wrap(~type) +
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1),
        
        #strip.background = element_blank(),
        
        panel.border = element_rect(colour = "black", fill = NA),
        
        strip.text = element_text(size = 10)) 

ggplotly(p)
  
```

### Shortcut

```{r improvements in word record - shortcut}


p <- records %>% 
  
  filter(shortcut == "Yes") %>% 
  
  group_by(track, type) %>% 
  
  summarise(change = max(time) - min(time),
         
         pct_change = change/max(time)) %>% 
  
  mutate(total = 1 -pct_change) %>% 
  
  ungroup() %>% 
  
  select(-change) %>% 
  
  mutate(track = fct_reorder(track, pct_change, .desc = FALSE)) %>% 
  
  ggplot(aes(pct_change, track, fill = type)) +
  
   geom_col()+
  
  #geom_col(aes(x = total, track)
  
  scale_x_continuous(labels = percent,
                     expand = c(0,0)) +
  
    facet_wrap(~type) +
  
  theme(axis.text.x = element_text(size = 7.5,angle = 45, vjust = 1, hjust=1),
        
        #strip.background = element_blank(),
        
        panel.border = element_rect(colour = "black", fill = NA),
        
        strip.text = element_text(size = 10)) 

ggplotly(p)
  
```

Players{data-icon="fa-user-plus"}
=======================================================================

Column
-----------------------------------------------------------------------

### Players With the Most World Records

```{r players with the most world records}

p <- drivers %>% 
  
  distinct(player, .keep_all = TRUE) %>% 
  
  slice_max(total, n=20) %>% 
  
  mutate(player = fct_reorder(player, total)) %>% 
  
  ggplot(aes(total, player)) +
  
   geom_col(fill = "turquoise3")+
  
  scale_x_continuous(expand = c(0,0)) +
  
  labs(x = "Total Records")
  
ggplotly(p)

```

Column
-----------------------------------------------------------------------

### Players With the Most Current World Records

```{r  players with the most current world records}

p <- records %>% 
  
  group_by(track) %>% 
  
  filter(time == min(time)) %>% 
  
  ungroup() %>% 
  
  count(player, sort = TRUE) %>% 
  
  mutate(player = fct_reorder(player, n, .desc = TRUE)) %>% 
  
  ggplot(aes(player, n)) +
  
  geom_col(fill = "tomato2") +
  
  labs(y = "Number of current records")

ggplotly(p)

```
