---
title: "Ask a manager survey - modelling"
author: "Antony"
date: "5/30/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(plotly)
library(lubridate)
library(ggthemes)
library(magrittr)
library(janitor)
library(ggsflabel)

options(scipen = 99)
theme_set(theme_light())

```

# Load the weekly Data

```{r Load}

#tt <- tt_load("2021-05-18")

data_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv')

```

# Helper Functions

```{r functions}

summarize_salary <- function(tbl){
  
  tbl %>% 
    
    summarize(n = n(),
              
              median_salary = median(annual_salary)) %>% 
    
    arrange(desc(n))
}


is_outlier <- function(x) {
  
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}


```

# Data Wrangling

```{r data wrangling}

skimr::skim(data_raw)

## Changing Structures

survey <- data_raw %>% 
  
  mutate(timestamp = mdy_hms(timestamp),
         
         age_category = fct_relevel(fct_reorder(how_old_are_you, parse_number(how_old_are_you)), "under 18"),
         
         overall_years_of_professional_experience = str_replace(overall_years_of_professional_experience, " - ", "-"),
         
         years_of_experience_in_field = str_replace(years_of_experience_in_field, " - ", "-"),
         
         overall_years_of_professional_experience = fct_reorder(overall_years_of_professional_experience, parse_number(overall_years_of_professional_experience)),
         
         years_of_experience_in_field = fct_reorder(years_of_experience_in_field , parse_number(years_of_experience_in_field )),
         
         gender = fct_collapse(gender, "Other or prefer not to answer" = c("Other or prefer not to answer", "Prefer not to answer")),
         
         race = coalesce(race, "Other")
         )
  

## Standardizing Industries

library(readx)
library(stringdist)

industries <- readxl::read_xlsx("2021_05_18_ask_a_manager_survey/industries.xlsx")

survey %<>% 
  
  mutate(industry_std = industries$Industry[amatch(industry, industries$Industry, maxDist = 3)],
         
         industry_std = ifelse(is.na(industry_std), "Other",industry_std)
         ) 



## USD Records

survey_usd <- survey %>% 
  
  filter(currency == "USD") %>% 
  
  #filter(annual_salary > quantile(annual_salary, .005),annual_salary <= quantile(annual_salary, .9995)) %>% 
  
  filter(annual_salary >= 5000,
         annual_salary <= 2000000) %>% 
  
  mutate(state = str_remove(state, ", .*")) %>% 
  
  mutate(state_abb = state.abb[match(state, state.name)],
         
       state_abb = ifelse(state == "District of Columbia", "DC", state_abb)) 
  

```



# Statictical Modelling

### Distribution of Salaries in USD

```{r Distribution of Salaries}

survey_usd %>% 
  
  ggplot(aes(annual_salary)) +
  
  geom_histogram() +
  
  scale_x_log10(labels = dollar)


```

Looks like a lognormal distribution

### Distribution of Salaries by State in USD

```{r Distribution of salaries by state - boxplot}

survey_usd %>% 
  
  filter(!is.na(state)) %>% 
  
  mutate(state = fct_lump(state, 9 ),
         
         state = fct_reorder(state, annual_salary)) %>% 
  
  mutate(outlier = ifelse(is_outlier(annual_salary), job_title, NA_character_)) %>% 
  
  ggplot(aes(annual_salary, state, fill = state)) +
  
  geom_boxplot()+
  
  scale_x_log10(labels = dollar) +
  
  theme(legend.position = "none")

```

So much variation within states as in between states, hence we'll look at the median as opposed to distribution

## ANOVA

- Compares the variance within groups and between groups

- Null hypothesis is that there is no variation, and that the means are equal.


### Which of overall experience and experience in field explain more variation?

```{r Linear Model - experience}

survey_usd %>% 
  
  filter(!is.na(overall_years_of_professional_experience)) %>% 
  
  ggplot(aes(annual_salary, overall_years_of_professional_experience)) +
  
  geom_boxplot() +
  
  scale_x_log10()



lm(log2(annual_salary)~overall_years_of_professional_experience, data = survey_usd) %>%   #Log 2 makes the intercepts more interpretable (i.e if 2, this means that the category makes 2^2 as much as the base category )
  
  summary()

lm(log2(annual_salary)~years_of_experience_in_field, data = survey_usd) %>% 
  
  summary()

```

* Overall Experience explains 5.8% of the overall variation
* Experience in field explains 10% of the variation
* Hence, experience in field might be a better predictor
* Note: We use log scale as that when the dependent variable (annual salary) is bell-curved

### What variations are explained by the different categories

```{r ANOVA}

library(broom)

survey_usd %>% 
  
  mutate(job_title = fct_lump(job_title, 30),
         
         state = fct_lump(state, 30)
         
         ) %>% 
  
  lm(log2(annual_salary) ~ job_title + 
       state + years_of_experience_in_field +
       gender + 
       race + 
       highest_level_of_education_completed +
       age_category+ 
       industry_std, 
     
     data = .) %>% 
  
  anova() %>% 
  
  broom::tidy() %>% 
  
  mutate(pct_variation =sumsq / sum(sumsq)) %>% 
  
  arrange(desc(pct_variation))


```

# Machine Learning

### Preprocessing

```{r machine learning preprocessing}

library(tidymodels)

set.seed(2021)

survey_usd_split <- initial_split(survey_usd)

survey_usd_training <- training(survey_usd_split)

survey_usd_testing <- testing(survey_usd_split)

skimr::skim(survey_usd)

rec <- survey_usd_training %>% 
  
  recipe(annual_salary ~ job_title + 
       state + years_of_experience_in_field +
       gender + 
       race + 
       highest_level_of_education_completed +
       age_category+ 
       industry_std) %>% 
  
  step_mutate_at(c("job_title", "state","highest_level_of_education_completed",  "industry_std"), fn = as.factor ) %>% # Changing to factor
  
  step_unknown(job_title, state, gender,highest_level_of_education_completed ) %>%  # Replacing NAs

  step_log(annual_salary, base = 2) %>% # Annual Salary to Log
  
  step_other(job_title, industry_std, state, threshold = tune()) %>%   # Collapsing Factor levels

  step_dummy(all_nominal()) #%>%  # Creating Dummy Variables for Factors
  
  #prep(training = survey_usd_training)
  
  
  
#training_prepared <- bake(rec, new_data = survey_usd_training)


```

### Tuning

```{r Tuning}

# V fold Cross Validation

training_cv <- vfold_cv(survey_usd_training)

# Tuning

threshold_grid <- c(".0001", ".0005", ".001", ".005", ".01", ".05")

linear_model_cv_tune_threshold <- linear_reg() %>% 
  
  set_engine("lm") %>% 
  
  tune_grid(rec,training_cv, threshold_grid)

linear_model_cv_tune_threshold %>% 
  
  collect_metrics() %>% 
  
  ggplot(aes(threshold, mean, color = .metric)) +
  
  geom_line()

```

Threshold best at lower values

### Cross Validation Final

```{r cross validation}

## Updating the recipe with a threhsold

rec_with_threshold <- rec %>% 
  
  finalize_recipe(list(threshold = .001))
  
linear_model_cv <- linear_reg() %>% 
  
  set_engine("lm") %>% 
  
  fit_resamples(rec_with_threshold, training_cv)

linear_model_cv$.metrics

linear_model_cv %>% 
  
  collect_metrics(summarize = FALSE)


linear_model_cv %>% 
  
  collect_metrics(summarize = FALSE) %>% 
  
  group_by(.metric) %>% 
  
  mutate(avg = mean(.estimate)) %>% 
  
  ungroup() %>% 
  
  ggplot(aes(id, .estimate, color = .metric, group = .metric)) +
  
  geom_line() +
  
  geom_hline(aes(yintercept = avg, color = .metric))

show_best(linear_model_cv, metric = "rmse")
```

### Comparing the Linear Model with Random Forest

```{r random forest}




```

