---
title: "US Salary - Ask a Manager Survey - US"
author: "Antony Rono"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(tidytuesdayR)
library(scales)
library(plotly)
library(lubridate)
library(ggthemes)
library(magrittr)
library(janitor)

options(scipen = 99)
theme_set(theme_light())

```

```{r Load, include=FALSE}

#tt <- tt_load("2021-05-18")

data_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv')

```

```{r functions, include=FALSE}

summarize_salary <- function(tbl){
  
  tbl %>% 
    
    summarize(n = n(),
              
              median_salary = median(annual_salary)) %>% 
    
    arrange(desc(n))
}


is_outlier <- function(x) {
  
  return(x < quantile(x, 0.25) - 1.5 * IQR(x) | x > quantile(x, 0.75) + 1.5 * IQR(x))
}



plot_categorical_bar <- function(data, col, n_level = 15, reorder = TRUE){
  
  lumped_data <- data %>% 
    
    filter(!is.na({{col}})) %>% 
    
    mutate({{col}} := fct_lump({{col}}, n_level ))
  
  if(reorder){
    
    lumped_data %<>%
      
      mutate({{col}} := fct_reorder({{col}}, annual_salary))
    
  }
          
  lumped_data %>% 
      
    mutate(outlier = ifelse(is_outlier(annual_salary), job_title, NA_character_)) %>% 
    
    group_by({{col}}) %>% 
    
    summarize_salary() %>% 
    
    ggplot(aes(median_salary, {{col}}, fill = {{col}})) +
    
    geom_col() +
    
    scale_x_continuous(labels = dollar, expand = c(0,0)) +
    
    theme(legend.position = "none") +
    
    labs(x = "Median Salary")
  
  
}

```

```{r data wrangling, include=FALSE}

skimr::skim(data_raw)

## Changing Structures

survey <- data_raw %>% 
  
  mutate(timestamp = mdy_hms(timestamp),
         
         age_category = fct_relevel(fct_reorder(how_old_are_you, parse_number(how_old_are_you)), "under 18"),
         
         overall_years_of_professional_experience = str_replace(overall_years_of_professional_experience, " - ", "-"),
         
         years_of_experience_in_field = str_replace(years_of_experience_in_field, " - ", "-"),
         
         overall_years_of_professional_experience = fct_reorder(overall_years_of_professional_experience, parse_number(overall_years_of_professional_experience)),
         
         years_of_experience_in_field = fct_reorder(years_of_experience_in_field , parse_number(years_of_experience_in_field )),
         
         gender = fct_collapse(gender, "Other or prefer not to answer" = c("Other or prefer not to answer", "Prefer not to answer")),
         
         race = coalesce(race, "Other")
         )
  

## Standardizing Industries

library(readxl)
library(stringdist)

industries <- readxl::read_xlsx("industries.xlsx")

survey %<>% 
  
  mutate(industry_std = industries$Industry[amatch(industry, industries$Industry, maxDist = 3)],
         
         industry_std = ifelse(is.na(industry_std), "Other",industry_std)
         ) 



## USD Records

survey_usd <- survey %>% 
  
  filter(currency == "USD") %>% 
  
  #filter(annual_salary > quantile(annual_salary, .005),annual_salary <= quantile(annual_salary, .9995)) %>% 
  
  filter(annual_salary >= 5000,
         annual_salary <= 2000000) %>% 
  
  mutate(state = str_remove(state, ", .*")) %>% 
  
  mutate(state_abb = state.abb[match(state, state.name)],
         
       state_abb = ifelse(state == "District of Columbia", "DC", state_abb)) 
  

```

Geographical Breakdown {data-icon="fa-map-marker-alt"}
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### Median Salary by State

```{r High Charter Map}

library(highcharter)

#x <- get_data_from_map(download_map_data("https://code.highcharts.com/mapdata/countries/us/us-all.js"))

survey_usd_median_salary_states <- survey_usd %>% 
  
  filter(!is.na(state)) %>% 
  
  mutate(state = fct_lump(state, 15 ),
         
         state = fct_reorder(state, annual_salary)) %>% 
  
  mutate(outlier = ifelse(is_outlier(annual_salary), job_title, NA_character_)) %>% 
  
  group_by(state_abb) %>% 
  
  summarize_salary() 


hcmap(
  
   map = "countries/us/us-all.js", # high resolution world map
  
   data = survey_usd_median_salary_states %>% rename(`hc-a2` = state_abb) , # name of dataset
  
   joinBy = "hc-a2",
  
   value = "median_salary",
  
   showInLegend = FALSE, # hide legend
  
   nullColor = "#FFFFFF",
  
   download_map_data = TRUE,
   
   dataLabels = list(enabled = TRUE, format = '{point.name}')

   ) %>% 
  
  #hc_colorAxis(minColor = "#FFFFCC", maxColor = "blue") %>% 
   
    hc_mapNavigation(enabled = FALSE) %>%
  
   hc_legend("none")

```

### Breakdown as Bar Chart

```{r Distribution of Median salaries by state}

p <- plot_categorical_bar(survey_usd, state, 15)

ggplotly(p, tooltip = c("x", "y"))

```

Column
-----------------------------------------------------------------------

### Relationship between cost of living and median salary

```{r Cost of living by state}

cost_of_Living <- read_csv("cost_of_living.csv")

cost_of_Living %<>% 
  
  clean_names() 

p <- survey_usd %>% 
  
  group_by(state) %>% 
  
  summarize_salary() %>% 
  
  ungroup() %>% 
  
  filter(!is.na(state)) %>% 
  
  left_join(cost_of_Living, by = "state") %>%
  
  filter(!is.na(cost_index)) %>% 
  
  ggplot(aes(median_salary, cost_index, color = state))+
  
  geom_point() +
  
  scale_x_continuous(labels = dollar)+
  
  theme(legend.position = "none")


ggplotly(p)


```

Salary by Profession {data-icon="fa-chart-bar"}
=======================================================================

Column
-----------------------------------------------------------------------

### Distribution of Salaries by Industry

```{r salaries by industry}

p <- plot_categorical_bar(survey_usd, industry_std, 15)
         
ggplotly(p, tooltip = c("x", "y"))
```


### Distribution of Salaries by Job Title

```{r salaries by job title}

p <- plot_categorical_bar(survey_usd, job_title, 15)

ggplotly(p, tooltip = c("x", "y"))

```

Column
-----------------------------------------------------------------------

### Distribution of Salaries by Overall Year of Experience

```{r salaries by overall experience}

p <- plot_categorical_bar(survey_usd, overall_years_of_professional_experience,  15, reorder = FALSE)

ggplotly(p, tooltip = c("x", "y"))
```

### Distribution of Salaries by Year of experience in field

```{r salaries by year of experience}

p <- plot_categorical_bar(survey_usd, years_of_experience_in_field,  15, reorder = FALSE)

ggplotly(p, tooltip = c("x", "y"))
```


Salary by Demography {data-icon="fa-chart-line"}
=======================================================================

Column
-----------------------------------------------------------------------

### Distribution of Salaries by Gender

```{r salaries by gender}

p <- plot_categorical_bar(survey_usd, gender,  15)

ggplotly(p, tooltip = c("x", "y"))

```

### Distribution of Salaries by Race

```{r salaries by race}

p <- plot_categorical_bar(survey_usd, race, n = 10)

ggplotly(p, tooltip = c("x", "y"))

```

Column
-----------------------------------------------------------------------

### Distribution of Salaries by Education Level

```{r salaries by education level}

p <- plot_categorical_bar(survey_usd, highest_level_of_education_completed)

ggplotly(p, tooltip = c("x", "y"))

```

### Distribution of Salaries by Age Category

```{r salaries by age}

p <- plot_categorical_bar(survey_usd, age_category, reorder = FALSE)

ggplotly(p, tooltip = c("x", "y"))

```


