---
title: "Netflix Titles"
author: "Antony Rono"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(flexdashboard)
library(tidyverse)
library(tidytuesdayR)
library(plotly)
library(janitor)
library(magrittr)
library(lubridate)
library(scales)
library(ggsci)


theme_set(theme_light())

```

```{r Load data, include=FALSE}

# tt <- tt_load("2021-04-20")
# 
# data <- tt$netflix_titles

data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv')

netflix <- data %>%
  
  clean_names() %>% 
  
  separate(duration, c("duration", "duration_units"), sep = " ", convert = TRUE) %>% 
  
  mutate(decade = 10*(release_year %/% 10))

skimr::skim(netflix)
```

```{r function to summarize movies, include=FALSE}

summarize_titles <- function(data){
  
  data %>% 
    
      summarize(n = n(),
            
            median_duration = median(duration),
            
            median_year = median(release_year),
            
            avg_duration = mean(duration),
            
            ) %>% 
    
    arrange(desc(n))
    
    
}
```


```{r function to summarize by count and percent, include=FALSE}

summarize_by_count_and_percent <- function(x, y){
  
  netflix %>% 
    
    separate_rows({{x}}, sep = ",") %>% 
    
    separate_rows({{y}}, sep = ",") %>% 
    
    mutate(
      
      {{x}} := str_trim({{x}}),
      
      {{y}} := str_trim({{y}})
      
    ) %>% 
    
    filter(!is.na({{x}}), !is.na({{y}})) %>% 
    
    mutate({{x}} := fct_lump_n({{x}},10),
           
           {{y}} := fct_lump_n({{y}},15)
    ) %>% 
    
    count({{x}}, {{y}}, name = "count", sort = TRUE) %>% 
    
    group_by({{x}}) %>% 
    
    mutate(pct = count/sum(count)) %>% 
    
    ungroup() %>% 
    
    mutate({{x}} := fct_reorder({{x}}, count, sum, .desc = FALSE),
           
           {{y}} := fct_reorder({{y}}, count, sum, .desc = FALSE)
    ) 
}

```

Overall {data-icon="fa-chart-line"}
=======================================================================

Row {data-height=250}
-----------------------------------------------------------------------
### Total Releases
```{r count of released}

valueBox(value = comma(nrow(data)), caption = "Total Releases", icon = "fa-user-plus", color = "green")

```

### Number of Countries Produced

```{r number of countries}

countries <- netflix %>% 
  
  separate_rows(country, sep = ",") %>% 
  
  mutate(country = str_trim(country)) %>% 
  
  distinct(country) %>% 
  
  nrow %>% 
  
  comma
  
gauge(countries, min = 0, max = 195, gaugeSectors(success = c(131, 195), warning = c(66,130), danger = c(0, 65)))

```


### Percentage of Movies and TV Shows

```{r proportion of movies and tv shows}

netflix %>% 
  
  count(type) %>% 
  
  plot_ly(labels = ~ type, values = ~ n) %>% 
  
  add_pie(hole = 0.6)

```

Row {data-height=750}
-----------------------------------------------------------------------

### Overall Trend in Productions in the Last Two Decades

```{r production by year  - stacked bar chart}

type_year_count_df <- netflix %>% 
  
    count(type, release_year) %>% 
  
    filter(release_year > 2000)
  
p <- type_year_count_df %>% 
  
  ggplot(aes(release_year, n)) +
  
  geom_col(aes(fill = type)) +
  
  geom_line(data = type_year_count_df %>% filter(type == "Movie"), color = "red1",  size = 1) +
  
  geom_line(data = type_year_count_df %>% filter(type == "TV Show"), color = "cyan4",  size = 1) +
  
  scale_y_continuous(expand = c(0,0)) +
  
  scale_x_continuous(expand = c(0,0), n.breaks = 11) +
  
  labs(y = "Count")


ggplotly(p)
```

### Relative Trend in TV Shows and Movies

```{r}
p <- netflix %>% 
  
  count(type, release_year) %>%
  
  group_by(release_year) %>% 
  
  pivot_wider(names_from = type, values_from = n, values_fill = 0) %>% 
  
  clean_names() %>% 
  
  mutate(pct_tv_shows = tv_show/sum(movie, tv_show),
         
         pct_movies = 1 - pct_tv_shows) %>% 
  
  
  pivot_longer(cols = starts_with("pct"), names_to = "type", values_to = "pct") %>% 
  
  mutate(type = ifelse(str_detect(type, "movies"), "Movie", "TV Show")) %>% 
  
  ggplot(aes(release_year, pct, color = type))+
  
  geom_point() +
  
  geom_line(arrow = arrow(angle = 15, ends = "last", type = "closed")) +
  
    scale_y_continuous(labels = percent, expand = c(0,0), limits = c(0,1.05)) +
  
   scale_x_continuous(expand = c(0,0), n.breaks = 15, limits = c(1925, 2022)) +
  
  labs(y = "Proportion of all releases")

ggplotly(p)

```

Country {data-orientation=columns data-icon="fa-location-arrow"}
=======================================================================

Column
-----------------------------------------------------------------------

### Total Number of Productions by Country

```{r total production by countr - worldmap}

library(highcharter)
library(maps)

prod_country <- netflix %>% 
  
  separate_rows(country, sep = ",") %>% 
  
  mutate(country = str_trim(country)) %>% 
  
  filter(str_detect(country, "")) %>% 
  
  count(country) %>% 
  
  mutate(country = case_when(
    
    str_detect(country, "Germany") ~ "Germany",
    
    country == "United Kingdom" ~ "UK",
    
    country == "Soviet Union" ~ "Russia",
    
    country == "Vatican City" ~ "Vatican",
    
    TRUE ~ country
    
  )) %>% 
  
  mutate(`iso-a3`= coalesce(iso3166$a3[match(country,iso3166$sovereignty)], iso3166$a3[match(country , iso3166$ISOname)])) 

hcmap(
  
   map = "custom/world-highres3", # high resolution world map
  
   data = prod_country, # name of dataset
  
   joinBy = "iso-a3",
  
   value = "n",
  
   showInLegend = FALSE, # hide legend
  
   nullColor = "#DADADA",
  
   download_map_data = TRUE
) %>% 
   
    hc_mapNavigation(enabled = FALSE) %>%
  
   hc_legend("none") #%>%
  
   #hc_title(text = "Production By Country") # title
 


```

Column
-----------------------------------------------------------------------

### Breakdown by Country and Type 

```{r production by country and type-worldmap }

p <- netflix %>% 
  
  separate_rows(country, sep = ",") %>% 
  
  mutate(country = str_trim(country)) %>% 
  
  filter(!is.na(country)) %>% 
  
  count(type, country = fct_lump(country, 15), sort = TRUE, name = "count") %>% 
  
  mutate(country = fct_reorder(country, count, sum),
         
         ) %>% 
  
  ggplot(aes(country, count, fill = type)) +
  
  geom_col() +
  
  scale_y_continuous(expand = c(0,0)) +
  
  coord_flip() +
  
  labs(y = "Count")

ggplotly(p)

```

### Breakdown By Ratings

```{r production by country and ratings}

p <- summarize_by_count_and_percent(country, rating) %>% 

  ggplot(aes(country, pct, fill = rating))+
  
  geom_col() +
  
  scale_y_continuous(expand = c(0,0), labels = percent) +
    
  coord_flip() +
  
  labs(y = "proportion of all releases")

ggplotly(p)

```

Genre {data-orientation=columns data-icon="fa-code"}
=======================================================================

Column
-----------------------------------------------------------------------

### Breakdown by Type

```{r production by genre}

p <- netflix %>% 
  
  separate_rows(listed_in, sep = ",") %>% 
  
  mutate(listed_in = str_trim(listed_in)) %>% 
  
  count(type, listed_in) %>% 
  
  mutate(listed_in = fct_reorder(listed_in, n, sum)) %>% 
  
  filter(!is.na(listed_in)) %>% 
  
  slice_max(n, n =30) %>% 
  
  ggplot(aes(listed_in, n, fill = type)) +
  
  geom_col() +
  
  scale_y_continuous(expand = c(0,0)) +
  
  coord_flip() +
  
  labs(y = "Count",
       
       x = "Genre") +
  
  facet_wrap(~type, scales = "free", ncol = 2) +
  
  theme(legend.position = "none") 


ggplotly(p)
```

Column
-----------------------------------------------------------------------

### Breakdown By Rating

```{r production by genre and ratings}

 p <- summarize_by_count_and_percent(listed_in, rating) %>% 

  ggplot(aes(listed_in, pct, fill = rating))+
  
  geom_col() +
  
  scale_y_continuous(expand = c(0,0), labels = percent) +
    
  coord_flip() +
  
  labs(y = "proportion of all ratings",
       
       x = "genre")

ggplotly(p)

```


Duration {data-icon="fa-calendar-check"}
=======================================================================

Row
-----------------------------------------------------------------------

### Trend of Length of Movies Overtime
```{r duration of movies over the years}

p <- ggplot() +
  
  geom_boxplot(data = netflix %>% filter(type == "Movie") , aes(decade, duration,group = decade), outlier.colour = "red", outlier.shape = 1) +
  
  geom_line(data = netflix %>% filter(type == "Movie") %>%  group_by(decade) %>% summarise(avg_length = median(duration)) %>% ungroup(), aes(decade, avg_length),size = 1, color = "darkorchid4")+
  
  scale_x_continuous(n.breaks = 15) 
  
ggplotly(p)
  

```

Row
-----------------------------------------------------------------------

### Breakdown By Country

```{r duration of movies by country}


p <- netflix %>% 
  
  separate_rows(country, sep = ",") %>% 
  
  mutate(country = str_trim(country)) %>% 
  
  group_by(type, country = fct_lump(country, 20)) %>% 
  
  summarize_titles() %>% 

  filter(type == "Movie") %>% 
  
  mutate(country = fct_reorder(country, median_duration)) %>% 
  
  filter(!is.na(country)) %>% 
  
  ggplot(aes(country, median_duration, fill = country)) +
  
  geom_col() +
  
  scale_y_continuous(expand = c(0,0)) +
  
  theme(legend.position = "none") +
  
  coord_flip() +
  
  labs(y = "Duration (minutes)",
       
       x = "Genre") 
  
ggplotly(p)
```

### Breakdown by Genre

```{r duration of movies by genre}

p <- netflix %>% 
  
  separate_rows(listed_in, sep = ",") %>% 
  
  mutate(listed_in = str_trim(listed_in)) %>% 
  
  group_by(type, listed_in) %>% 
  
  summarize_titles() %>% 

  filter(type == "Movie") %>% 
  
  filter(listed_in != "Movies") %>% 
  
  mutate(listed_in = fct_reorder(listed_in, median_duration)) %>% 
  
  filter(!is.na(listed_in)) %>% 
  
  ggplot(aes(listed_in, median_duration, fill = listed_in)) +
  
  geom_col() +
  
  scale_y_continuous(expand = c(0,0)) +
  
  theme(legend.position = "none") +
  
  coord_flip() +
  
  labs(y = "Duration (minutes)",
       
       x = "Genre") 
  
ggplotly(p)

```

Ratings {data-icon="fa-thumbs-up"}
=======================================================================

Row
-----------------------------------------------------------------------

### Trend of Ratings Overtime

```{r ratings overtime trend}

p <- netflix %>% 
  
  filter(!is.na(rating)) %>% 
  
  mutate(rating = fct_lump_n(rating,5)) %>% 
  
  count(rating, decade, name = "count", sort = TRUE) %>% 

  mutate(rating = fct_reorder(rating, count, sum, .desc = TRUE)) %>% 
  
  ggplot(aes(decade, count, color = rating))+
  
  geom_point()+
  
  geom_line() +
  
  scale_x_continuous(n.breaks = 15) 

ggplotly(p)


```

### Ratings By Country

```{r ratings by country}

p <- summarize_by_count_and_percent(rating, country) %>% 

  ggplot(aes(rating, pct, fill = country))+
  
  geom_col() +
    
  coord_flip() +
  
  scale_y_continuous(labels = percent)+
  
  #scale_fill_discrete(palette = "Spectral")+
  
  labs(y = "proportion of all releases")

ggplotly(p)

```


Row
-----------------------------------------------------------------------

### Countries that have mature content (R/TV-MA)

```{r countris by mature content}
p <- netflix %>% 
  
  separate_rows(country, sep = ",") %>% 
  
  mutate(country=str_trim(country)) %>% 
  
  filter(!is.na(rating), !is.na(country)) %>%
  
  group_by(type, country = fct_lump(country, 10)) %>% 
  
  summarize(n_mature = sum(rating %in% c("R", "TV-MA", "NC-17")),
            
            n = n(),
            
            .groups = "drop"
            
            ) %>% 
  
  mutate(pct_mature = n_mature / n,
         
         conf_low = qbeta(0.025, n_mature + 0.5, n - n_mature + 0.5),
         
         conf_high = qbeta(0.975, n_mature + 0.5, n - n_mature + 0.5),
         
         ) %>% 
  
  ggplot(aes(pct_mature, country, color = type))+
  
  geom_point(aes(size = n)) +
  
  geom_errorbar(aes(xmin = conf_low, xmax =conf_high ), width = 0.5) +
  
  scale_x_continuous(labels = percent, n.breaks = 4) +
  
  labs(x = "% of releases that are Mature (R/TV-MA/NC-17 rated)",
       
       size = "No. of releases")

ggplotly(p)

```


### Rating By Genre

```{r ratings by genre}

p <- summarize_by_count_and_percent(rating, listed_in) %>% 

  ggplot(aes(rating, pct, fill = listed_in))+
  
  geom_col() +
    
  coord_flip() +
  
  scale_y_continuous(labels = percent)+
  
  #scale_fill_discrete(palette = "Spectral")+
  
  labs(y = "proportion of all releases",
       
       fill = "Genre")

ggplotly(p)

```



Word Frequency {data-orientation=columns data-icon="fa-globe"}
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### Frequency of words in title -Wordcloud

```{r frequency of words in title - wordcloud}
library(tidytext)
library(wordcloud2)
library(webshot)
library(htmlwidgets)

word_title_freq <- netflix %>% 
  
  unnest_tokens(word, title, drop = FALSE) %>% 
  
  anti_join(stop_words, by = "word") %>% 
  
  filter(str_length(word) >2) %>% 
  
  count(word) %>% 
  
  arrange(desc(n)) %>% 
  
  rename(freq = n) %>% 
  
  filter(freq > 1)
  
  #slice_max(n, n = 70)


graph <- wordcloud2(word_title_freq, size=0.8, terrain.colors(length(word_title_freq$word)), shape = "circle", backgroundColor = "black")

#webshot::install_phantomjs()

# save it in html

saveWidget(graph,"netflix_titles_wordcloud.html",selfcontained = F)

# and in png or pdf
webshot("netflix_titles_wordcloud.html","netflix_titles_wordcloud.png", delay =8, vwidth = 480, vheight=480)

graph

```
![](netflix_titles_wordcloud.html)

### Most Common words in titles for each type

```{r most common words in description}
library(snakecase)
library(tidylo)

words_unnested <- netflix %>%

  unnest_tokens(word, description, drop = FALSE) %>%

  anti_join(stop_words, by = "word") %>%

  filter(str_length(word) >2)


p <- words_unnested %>%

  count(type, word) %>%

  bind_log_odds(type, word, n) %>%

  arrange(desc(log_odds_weighted)) %>%

  group_by(type) %>%

  slice_max(log_odds_weighted, n=15) %>%

  ungroup() %>%

  mutate(word = fct_reorder(word, log_odds_weighted)) %>%

  ggplot(aes(log_odds_weighted, word, fill = type)) +

  geom_col() +

  scale_x_continuous(expand = c(0,0))+

  theme(legend.position = "none")+

  facet_wrap(~type, scales = "free_y") +

  labs(x = "Log odds ratio, weighted by uninformative Dirichlet prior")

ggplotly(p)

```

Column
-----------------------------------------------------------------------

### Cluster of words that tend to appear together in movie/tv description

```{r clustering of words}

library(widyr)
library(tidygraph)
library(ggraph)

words_unnested <- netflix %>% 
  
  unnest_tokens(word, description, drop = FALSE) %>% 
  
  anti_join(stop_words, by = "word") %>% 
  
  filter(str_length(word) >2)
  
  
words_unnested %>% 
  
  distinct(type, title, word ) %>% 
  
  add_count(word, name = "word_total") %>% 
  
  arrange(desc(word_total)) %>% 
  
  filter(word_total >=40) %>% 
  
  pairwise_cor(word, title, sort = TRUE) %>% 
  
  filter(correlation > 0.1) %>% 
  
  igraph::graph_from_data_frame() %>% 
  
  #ggraph(layout = "igraph", algorithm = 'kk') +
  
  ggraph(layout = "fr")+
  
  geom_edge_link(aes(alpha = correlation)) +
  
  geom_node_point(col = "red") +
  
  #geom_node_text(aes(label = name), hjust = 1, vjust =1 , check_overlap = TRUE)

  geom_node_text(aes(label = name), repel = TRUE) +
  
  theme(legend.position = "none")+
  
  scale_color_brewer(palette = "Set1") +
  
  scale_fill_brewer(palette = "Set1")

```



