---
title: "US Broadband"
author: "Antony Rono"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(flexdashboard)
library(tidyverse)
library(tidytuesdayR)
library(scales)
library(zipcodeR)
library(janitor)
library(magrittr)
library(tigris)
library(ggthemes)
library(highcharter)


options(tigris_use_cache = TRUE)
options(scipen = 99)
theme_set(theme_light())
```

```{r Load, include=FALSE}

tt <- tt_load("2021-05-11")

broadband <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-11/broadband.csv')

broadband_zip <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-11/broadband_zip.csv')


```

```{r Cleaning Data, include=FALSE}

broadband_county <- broadband %>% 
  
  clean_names() %>% 
  
  mutate(state = state.name[match(st, state.abb)],
         
         state = ifelse(st == "DC", "District of Columbia", state),
         
         broadband_availability_per_fcc = parse_number(broadband_availability_per_fcc, na = "-"),
         
         broadband_usage = parse_number(broadband_usage, na = "-"),
         
         county_id = sprintf("%05d", county_id)
         
         ) %>% 
  
  mutate_at(c("county_name", "state"), str_trim)

skimr::skim(broadband)
  

```

```{r Adding County Population to broadband df, include=FALSE}

library(readxl)

population <- read_xlsx("US Population By County.xlsx", skip = 3)

population %<>% 
  
  rename(county_name = ...1) %>% 
  
  clean_names() %>% 
  
  filter(!is.na(census)) %>% 
  
  mutate(county_name = str_remove(county_name, "^[.]")) %>% 
  
  separate(county_name, c("county_name", "state"), sep = ",") %>% 
  
  mutate_at(c("county_name", "state"), str_trim) %>% 
  
  select(county_name, state, population_2017 = x2017, population_2019 = x2019) %>% 
  
  mutate(county_name = case_when(
    
    county_name == "Petersburg Borough" ~ "Petersburg Census Area",
    
    county_name == "Oglala Lakota County" ~ "Oglala County",
    
    TRUE ~ county_name
    
    
  ))
  


broadband_with_population <- broadband_county %>% 
  
  inner_join(population, by = c("county_name", "state")) %>% 
  
  mutate(county_state = paste0(str_remove(county_name, " County"),", ", state))
  
  
```

```{r Cleaning and Joining with zip code db, include=FALSE}

broadband_zip %<>% 
  
  clean_names()

skimr::skim(broadband_zip)

broadband_zip_joined <- broadband_zip %>% 
  
  mutate(postal_code = sprintf("%05d", postal_code)) %>% 
  
  left_join(zip_code_db, by = c("postal_code" = "zipcode")) %>% 
  
   mutate(state = state.name[match(st, state.abb)],
         
         state = ifelse(st == "DC", "District of Columbia", state)
   ) %>% 
  
  mutate(city_state = paste0(major_city,", ", state)) %>% 
  
  select(-county) 

```

Overall {data-icon="fa-map-marker-alt"}
=======================================================================


Row {data-height=250}
-----------------------------------------------------------------------

### Total Number of Counties

```{r count of counties}


valueBox(value = comma(nrow(broadband_county %>% distinct(county_id))), caption = "Number of Counties", icon = "fa-map-pin", color = "green")


```

### Number of Counties with more than 50% Broadband Acess

```{r valuebox of broadband access}

gauge(broadband_county %>% filter(broadband_availability_per_fcc > 0.5) %>%  nrow,
      min = 0, 
      max = nrow(broadband_county), 
      gaugeSectors(success = c(0.75*nrow(broadband_county), nrow(broadband_county)), 
                   warning = c(0.5*nrow(broadband_county),(0.75*nrow(broadband_county)-1)), 
                   danger = c(0, 0.5*nrow(broadband_county)))
      )

```

### Number of Counties with more than 25% Broadband Usage

```{r valuebox of broadband usage}

gauge(broadband_county %>% filter(broadband_usage > 0.25) %>%  nrow,
      min = 0, 
      max = nrow(broadband_county), 
      gaugeSectors(success = c(0.75*nrow(broadband_county), nrow(broadband_county)), 
                   warning = c(0.5*nrow(broadband_county),(0.75*nrow(broadband_county)-1)), 
                   danger = c(0, 0.5*nrow(broadband_county)))
      )

```


Row{data-height=750}
-----------------------------------------------------------------------

### Broadband Access in the Different Counties

```{r Broadband Access in Different Counties}

hcmap(
  
   map = "countries/us/us-all-all-highres", # high resolution world map
  
   data = broadband_with_population %>% rename(fips = county_id) , # name of dataset
  
   joinBy = "fips",
  
   value = "broadband_availability_per_fcc",
  
   showInLegend = FALSE, # hide legend
  
   nullColor = "#FFFFFF",
  
   download_map_data = TRUE,
   
   dataLabels = list(enabled = TRUE, format = '{point.name}')

   ) %>% 
  
  #hc_colorAxis(minColor = "#FFFFCC", maxColor = "blue") %>% 
   
    hc_mapNavigation(enabled = FALSE) %>%
  
   hc_legend("none")
```


### Broadband Usage in the Different Counties

```{r Broadband Usage in Different Counties}

hcmap(
  
   map = "countries/us/us-all-all-highres", # high resolution world map
  
   data = broadband_with_population %>% rename(fips = county_id) , # name of dataset
  
   joinBy = "fips",
  
   value = "broadband_usage",
  
   showInLegend = FALSE, # hide legend
  
   nullColor = "#FFFFFF",
  
   download_map_data = TRUE,
   
   dataLabels = list(enabled = TRUE, format = '{point.name}')

   ) %>% 
  
  #hc_colorAxis(minColor = "#FFFFCC", maxColor = "blue") %>% 
   
    hc_mapNavigation(enabled = FALSE) %>%
  
   hc_legend("none")

```


Relationships {data-orientation=columns data-icon="fa-chart-line"}
=======================================================================

Column
-----------------------------------------------------------------------

### Relationship Between Broadband Usage and Broadband Availability

```{r Broadband Usage by Broadband Availability}

p <- broadband_with_population %>% 
  
  ggplot(aes(broadband_availability_per_fcc,broadband_usage, color = county_state )) +
  
  geom_point() +
  
  geom_smooth(color = "blue", method = "loess")+
  
  theme(legend.position = "none") +
  
  labs(y = "Broadband Usage in 2019",
       
       x = "Broadband Availability in 2017")


p


```

Column
-----------------------------------------------------------------------

### Relationship Between Broadband Usage and Population Size

```{r Broadband Usage by Population Size}

p <- broadband_with_population %>% 
  
  ggplot(aes( broadband_usage,population_2019, color = county_state )) +
  
  geom_point() +
  
  geom_smooth(color = "blue", method = "lm")+
  
  theme(legend.position = "none") +
  
  scale_x_continuous(labels = percent) +
  
  scale_y_log10(labels = comma) +
  
  coord_flip() +
  
  expand_limits(y = 10000)+
  
  labs(x = "Broadband Usage in 2019",
       
       y = "Population in 2019 (Log Scale)")

p

```


### Relationship Between Broadband Usage and Median Income

```{r Broadband Usage by HH Income}

p <- broadband_zip_joined %>% 
  
  filter(!is.na(median_household_income)) %>% 
  
  ggplot(aes(median_household_income, broadband_usage , color = city_state )) +
  
  geom_point() +
  
  geom_smooth(color = "blue", method = "lm")+
  
  scale_x_log10(labels = comma) +
  
  scale_y_continuous(labels = percent, limits = c(0,1)) +
  
  theme(legend.position = "none") +
  
  labs(y = "Broadband Usage in 2020",
       
       x = "Median Household Income")

p

```


Column
-----------------------------------------------------------------------

### Relationship Between Broadband Usage and Population Density

```{r Broadband Usage by Population density}

p <- broadband_zip_joined %>% 
  
  filter(population > quantile(population, 0.2, na.rm = TRUE)) %>% 
  
  ggplot(aes(population_density, broadband_usage , color = city_state )) +
  
  geom_point() +
  
  geom_smooth(color = "blue", size =2, method = "lm")+
  
  scale_x_log10(labels = comma) +
  
  scale_y_continuous(labels = percent) +
  
  
  theme(legend.position = "none") +
  
  labs(y = "Broadband Usage in 2020",
       
       x = "Population Density (Log Scale)")

p

```

### Relationship Between Broadband Usage and Occupied Housing Units

```{r, Broadband Usage by Occupied Housing Units}

p <- broadband_zip_joined %>% 
  
  filter(!is.na(occupied_housing_units), occupied_housing_units > 0) %>% 
  
  ggplot(aes(occupied_housing_units, broadband_usage , color = city_state )) +
  
  geom_point() +
  
  geom_smooth(color = "blue", method = "lm")+
  
  scale_x_log10(labels = comma) +
  
  scale_y_continuous(labels = percent, limits = c(0,1)) +
  
  theme(legend.position = "none") +
  
  labs(y = "Broadband Usage in 2020",
       
       x = "Occupied Housing Units")

p

```


